Tropical Andes Functional diversity for plants and Frugivores

### 2024-05-16 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
remove unused packages, add figure path, add code to map without projection -- still doesn't work

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(ggspatial)
library(vegan)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

@@

@@

  # Calculate functional dispersion for grid
  fdis_grid <- TAGrid %>%
    st_join(data) %>%
    group_by(cellid) %>%
    summarize(sum_variable = sum(variable_name, na.rm = TRUE))

@@

  # Return a list of data
  list(FDplot = FDplot)
}
```

@@

# Plants
Test with small random sample as a subset to make it work before using all data
```{r}
all_plant_species <- row.names(plant_traits_df_final)
plant_random_subset <- sample(all_plant_species, 1000)
# objects that need to be subset
PAM_plant_site_final #matrix
plant_traits_df_final
```

>>>>

# Load required packages
```{r}
library(mFD)
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
library(rlang)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

@@

# Mapping without projection 
Because projecting is being difficult
```{r}
TropicalAndes_IUCNHabitat_Forest_latlong <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
Americas_latlong <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly_latlong <- Americas_latlong %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

```{r}
# Generate coordinates
subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

```{r}
# Create a grid over the area
create_grid <- function(data, cellsize) {
  st_make_grid(st_as_sfc(st_bbox(data)), cellsize = cellsize) %>%
    st_sf() %>%
    mutate(cellid = row_number())
}

# Aggregate functional dispersion within each grid cell
aggregate_to_grid <- function(grid, points, value_col) {
  points %>%
    st_join(grid) %>%
    group_by(cellid) %>%
    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
    st_set_geometry("geometry")
}

# Example usage with your data
resolution_meters <- 100000
variable_name <- "fdis_frugivore"

# Create grid
grid <- create_grid(spatial_fdis_frugivore, cellsize = resolution_meters)

# Aggregate data to grid
aggregated_data <- aggregate_to_grid(grid, spatial_fdis_frugivore, variable_name)

```{r}
FDisplot <-
    ggplot() +
    geom_sf(data = Americas_latlong, fill = "white")+
    geom_sf(data = TApoly_latlong, fill = "lightgrey", size = 0.1) +
    geom_point(data = spatial_fdis_frugivore , aes(fill = fdis_frugivore)) +
    scale_fill_gradient(low = "blue", high = "red") +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue")) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis")
FDisplot
```

@@

  # Calculate functional dispersion for grid
  fdis_grid <- TAGrid %>%
    st_join(data) %>%
    group_by(cellid) %>%
    summarize(
      sum_variable = sum(variable_name, na.rm = TRUE),
      mean_variable = mean(variable_name, na.rm = TRUE)
    )

@@

  # Return a list of data
  list(fdis_grid = fdis_grid,
        FDplot = FDplot)
}
```

@@

# Plants
Test with small random sample as a subset to make it work before using all data
```{r}
set.seed(100)
all_plant_species <- row.names(plant_traits_df_final)
plant_random_subset <- sample(all_plant_species, 1000)
# objects that need to be subset
PAM_plant_site_final_df <- as.dataframe(PAM_plant_site_final) #matrix
plant_traits_df_final
```
